---
title: "Reproducible Research Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load the package and data

Load the package.
```{r}
library(dplyr)
library(lattice)
```


Download and upzip data.

```{r eval = FALSE}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "activity.zip")
unzip("activity.zip")
```

Read the data. Transform the date column into date format.

```{r}
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
str(data)
summary(data)
head(data,6)
```


## What is mean total number of steps taken per day?

```{r}
steps_per_day <- data %>% group_by(date) %>% summarise_at(vars(steps), funs(sum(., na.rm = TRUE)))
hist(steps_per_day$steps, main = "Total number of steps taken per day")
steps_per_day_mean = mean(steps_per_day$steps)
steps_per_day_median = median(steps_per_day$steps)
```

**The mean of the total number of steps taken per day is `r steps_per_day_mean`, the median is `r steps_per_day_median`.**


## What is the average daily activity pattern?

```{r}
steps_per_interval <- data %>% group_by(interval) %>% summarise_at(vars(steps), funs(steps_avg = mean(., na.rm = TRUE)))
head(steps_per_interval)
with(steps_per_interval, plot(interval, steps_avg, type = "l"))
steps_per_interval_max <- max(steps_per_interval$steps_avg)
steps_per_interval_max_loc <- steps_per_interval$interval[which(steps_per_interval$steps_avg == steps_per_interval_max)]

```

**The 5-minute interval whose value is `r steps_per_interval_max_loc` has the maximum averaged number of steps.**  
**The maximum is `r steps_per_interval_max`.**


## Imputing missing values

```{r}
num_missing <- sum(is.na(data$steps))
```

**The total number of missing values is `r num_missing`. **  

Use the averaged number of steps in each 5-min interval to imput the missing data.  

```{r}
data2 <- data
missingRows <- data2[is.na(data2$steps), ]
imputedRows<- left_join(missingRows, steps_per_interval, by = "interval")
data2$steps[is.na(data2$steps)] <- imputedRows$steps_avg
```

data2 is the new dataset after imputation of missing values.

```{r}
steps_per_day2 <- data2 %>% group_by(date) %>% summarise_at(vars(steps), funs(sum(., na.rm = TRUE)))
hist(steps_per_day2$steps, main = "Total number of steps taken per day")
steps_per_day_mean2 = mean(steps_per_day2$steps)
steps_per_day_median2 = median(steps_per_day2$steps)
```

**After imputing the missing values, the mean of the total number of steps taken per day is `r steps_per_day_mean2`, the median is `r steps_per_day_median2`. The values differ from the estimates from the first part of the assignment. Imputing missing data reduces the frequency of having a small total number of steps in a day, and makes the histogram more symmetric. **


## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
data2 <- mutate(data2, day = weekdays(date))
whatdayis <- data.frame(day = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"), 
                        isWeekday = c("weekday", "weekday", "weekday", "weekday", "weekday", "weekend", "weekend"))
data2 <- left_join(data2, whatdayis, by = "day")
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
steps_per_interval2 <- data2 %>% group_by(isWeekday, interval) %>% summarise_at(vars(steps), funs(steps_avg = mean(., na.rm = TRUE)))
xyplot(steps_avg ~ interval | isWeekday, data = steps_per_interval2, type = "l", layout = c(1, 2))
```

